---
title: "Customer Behavior (E-commerce) Analysis"
author: "William L."
format: 
  html:
    self-contained: True
---

##### WD: setwd("/Users/williamlorenzo/Desktop/E-commerce Analysis in R")

## Background

Understanding customer behavior in e-commerce is crucial because it allows businesses to tailor their strategies to meet customer needs and preferences, ultimately driving sales and improving customer satisfaction.

## Loading data

```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
library(GGally)

data <- read.csv("E-commerce.csv")
```

#### Inspection of dataset structure

```{r}
str(data)
```

#### Summary of dataset

```{r}
summary(data)
```

## Data Cleaning

I will conduct missing values and outliers

#### Handle missing value

```{r}
data <- data %>%
  mutate(across(everything(), ~replace_na(.x, median(.x, na.rm = TRUE))))
```

#### Correct data types

```{r}
data$Age <- as.numeric(data$Age)
data$Annual.Income <- as.numeric(gsub("[^0-9]", "", data$Annual.Income))
data$Time.on.Site <- as.numeric(data$Time.on.Site)
```

#### Duplicate detection

```{r}
data <- data[!duplicated(data$Customer.ID), ]
```

#### Process of purchase history

Upon analyzing the process of purchase history, I have the ability to analyze how many purchases each customer has made. 

This is crucial for understanding customer behavior, segmenting customers based on their purchase frequency, and tailoring marketing strategies accordingly.
```{r}
data$Purchase.Count <- sapply(data$Purchase.History, function(x) length(strsplit(as.character(x), ",")[[1]]))

head(data)
```

## Descriptive Statistics

#### Summary statistics for numerical features

```{r}
summary(data$Age)
summary(data$Annual.Income)
summary(data$Time.on.Site)
```

#### Freq. dist. for categorical features

```{r}
table(data$Gender)
table(data$Location)
table(data$Product.Category)
```

#### Numerical Data using Histograms

In order to analyze numerical data, I have created two histograms using `ggplot()` to furhter visualize the distributions of `Age` and `Annual.Income` in my dataset.

```{r}
ggplot(data, aes(x = Age)) +
  geom_histogram(binwidth = 5,
                 fill = "blue",
                 color = "black") +
  ggtitle("Age Distribution")

ggplot(data, aes(x = Annual.Income)) +
  geom_histogram(binwidth = 10000, fill = "purple", color = "black") +
  ggtitle("Income Distribution")
```

#### Boxplot for Time on Site

Boxplot visualization to portray the distributions of `Gender` and `Location`.

```{r}
ggplot(data, aes(x = Gender)) +
  geom_bar(fill = "lavender") +
  ggtitle("Gender Distribution")

ggplot(data, aes(x = Location)) +
  geom_bar(fill = "lightblue") +
  ggtitle("Location Distribution")

```

#### Heatmap for Correlation Matrix

`heatmap()` visualizes the correlation matrix of the numeric variables

  - The strength and direction of linear relationships between pairs of numeric variables
  
  - Identify which variables are strongly correlated (positively or negatively)

```{r}
cor_matrix <- cor(data[, sapply(data, is.numeric)])
heatmap(cor_matrix, symm = T, col = heat.colors(256), main = "Correlation Heatmap")
```

## Data Visualization

#### Visualizing the Data to portray trends

I will begin pairing plots to depict relationships between all numerical features

```{r}
ggpairs(data[, sapply(data, is.numeric)], title = "Pair Plot of Numerical Features")
```

#### Comparison of distributions by gender

```{r}
ggplot(data, aes(x = Annual.Income, fill = Gender)) +
  geom_density(alpha = 0.5) +
  ggtitle("Income Distribution by Gender")
```

#### Time on Site v. Purchase Count

```{r}
ggplot(data, aes(x = Time.on.Site, y = Purchase.Count)) +
  geom_point(color = "lightblue") +
  geom_smooth(method = "lm", color = "blue") +
  ggtitle("Time on Site vs. Purchase Count")
```

